---
title: "RepData"
author: "Ben Meddeb"
date: "April 15, 2015"
output: html_document
keep_md: true
---


## Loading and preprocessing the data

Manually download and unzip [the Activity dataset](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) 

```{r  echo = TRUE}
library (plyr)
library(data.table)
library (xtable)
library(lattice)
Activity <- read.csv("activity.csv")
Activity <- data.table(Activity, keep.rownames=FALSE) 

```



## What is mean total number of steps taken per day?

We are going to use the 'plyr package to summarize the data ommiting the NAs

```{r  echo = TRUE}
nsd <- ddply (Activity, .(date), summarize, steps=sum(steps, na.rm=TRUE))
```
Draw a quick Histogram to summrize the data : 
```{r}
hist(nsd$steps, col='red', main ='Total number of steps taken per day', xlab = 'Mean Number of Steps', breaks = 10) 
rug(nsd$steps, col=2, lwd=2)
```


Display the number of steps summary per day : 
```{r}
xtable(summary(nsd), type="html")
```

Calculate the mean and median of the total number of recorded steps : 
```{r}
steps_mean <- mean(nsd$steps)
steps_median <- median(nsd$steps)
```

**The data indicates that the mean of the total number of steps per day is ** *`r steps_mean`* **with a median of** *`r steps_median`.*


## What is the average daily activity pattern?

To average the daily pattern we will use the plyr package again to summarize by the intervals recorded ommiting NAs: 

```{r}
adp <- ddply (Activity, .(interval), summarize, steps = mean(steps, na.rm=TRUE))
```

Display the summary : 

```{r}
xtable(summary(adp))
```

Quick plot

```{r}
plot(adp, type = "l", col = "red")
```


## Imputing missing values

Calculating the number of missing values : 

```{r}
missingValue <- sum(is.na(Activity))
```

The number of missing values is **`r missingValue`**

We have already calculated the average number of steps recorded per day and we stored the result in `adp` , So I am going to replace the NAs with the corresponding average and store the new data in `newActivity` 

```{r}
newActivity <- Activity
newActivity$steps[is.na(Activity$steps)] <- adp$steps[adp$interval == newActivity$interval[is.na(Activity$steps)]]
newActivity$steps[is.na(Activity$steps)] <- mean (Activity$steps, na.rm=TRUE)
```

Summary of the newly compiled data `newActivity` to compare it to the original data `Activity`, Just like before we will use the plyr package : 

``` {r}
newNsd  <- ddply(newActivity, .(date), summarize, steps=sum(steps))
xtable(summary(newNsd), type="html")
``` 


Histogram and summary of the new data 

```{r}
hist(newNsd$steps, col='red', main ='Total number of steps taken per day', xlab = 'Mean Number of Steps', breaks = 10) 
rug(nsd$steps, col=2, lwd=2)
```


## Are there differences in activity patterns between weekdays and weekends?

To answer the question we have to create a new factor weekend or weekday by adding a new column `weekdays` : 

```{r}
weekdays <- weekdays (as.Date (newActivity$date))
newActivity$dayType <- factor (weekdays, levels <- c ('weekday', 'weekend'))
newActivity$dayType[] <- 'weekday'
newActivity$dayType[weekdays %in% c ('Saturday', 'Sunday')] <- 'weekend'
```

plot the answer : 

```{r}
newAdp <- ddply (newActivity, .(interval, dayType), summarize, stepMean = mean(steps, na.rm=TRUE))
xyplot (stepMean ~ interval | dayType, data=newAdp, type='l', layout=c(1, 2))
```


# End 